---
output:
  pdf_document: default
  html_document: default
  word_document: default
---


```{r}
hData=read.csv('hotel_bookings.csv')
head(hData)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(pROC))
```
# Basic Analysis of dataset

```{r}
str(hData)

```
```{r}
table(is.na(hData))
colSums(is.na(hData))

```

# Analysis of Type of Hotel


```{r}
ggplot(data = hData,aes( x = hotel,y = prop.table(stat(count)),label = scales::percent(prop.table(stat(count))))) +
geom_bar(fill = 'steelblue') +geom_text(stat = "count", position = position_dodge(.9),vjust = -0.5,size = 3.8)+
scale_y_continuous(labels = scales::percent) +labs(title = "Bookings by Hotel Type",x = "Hotel Type",y = "Number of bookings")+geom_text(stat = "count", aes(label = ..count..), vjust = 5) 

```

"""

City Hotel   - 66%  - 79330
Resort Hotel - 34%  - 40060

People prefer City Hotel


"""

#Analysis of Is_cancelled



```{r}
ggplot(data = hData,aes( x = is_canceled,y = prop.table(stat(count)),
                         label = scales::percent(prop.table(stat(count))))) +
  geom_bar(fill = 'steelblue')  +geom_text(stat = "count", aes(label = ..count..), vjust = 5) +
  geom_text(stat = "count", position = position_dodge(.9),vjust = -0.5,size = 3)+
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Bookings Cancelled", x = "Is Cancelled",y = "Number of Bookings") 

```

"""

Bookings cancelled

Yes - 37%  - 44224
No  - 63%  - 75166


"""

 
#Analysis of Lead Time
```{r}
summary(hData$lead_time)
```
#Analysis of Arrival Date Year

```{r}
ggplot(data = hData,aes( x = arrival_date_year,y = prop.table(stat(count)),
                         label = scales::percent(prop.table(stat(count))))) +
  geom_bar(fill = 'steelblue') +  geom_text(stat = "count", position = position_dodge(.9),vjust = -0.5,size = 3)+
  scale_y_continuous(labels = scales::percent) +  labs(title = "Bookings by Year", x = "Year of Arrival Date",y = "Number  of bookings") +  geom_text(stat = "count", aes(label = ..count..), vjust = 5)
```
"""

Year 
2015  - 18%
2016  - 47%
2017  - 34%

"""
#Analysis of Arrival Month

```{r}
hData$arrival_date_month =  factor(hData$arrival_date_month, levels = month.name)

ggplot(data = hData,aes( x = arrival_date_month,y = prop.table(stat(count)),
label = scales::percent(prop.table(stat(count))))) +  geom_bar(fill = 'steelblue') + 
geom_text(stat = "count", position = position_dodge(1),vjust = 1, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent) +
labs(title = "Bookings by Month", x = "Month of Arrival Date",y = "Number of bookings")+ coord_flip() +
  geom_text(stat = "count", aes(label = ..count..), hjust = 5) 
   
 

```
"""

 Bookings are made highest in August and July
 Lowest in January

"""
# Analysis of Meal

```{r}
ggplot(data = hData,aes( x = meal,fill = meal,y = prop.table(stat(count)),
label = scales::percent(prop.table(stat(count))))) +  geom_bar() + 
geom_text(stat = "count", position = position_dodge(1),vjust = 1, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent) + coord_flip() +labs(title = "Bookings by Meal", x = "Meal type",y = "Number of bookings") +
  geom_text(stat = "count", aes(label = ..count..), hjust = 5) 

```
"""

Bookings made with Bed and Breakfast are 77%
Full Board meal bookings are low with 0.67%
"""


"""

As the hotels are located in Portugal Europe, the booking are mostly with European countries 
Highest is portugal with 48590 bookings

"""

# Analysis of Market Segment

```{r}

ggplot(data = hData,aes( x = market_segment,fill = market_segment,y = prop.table(stat(count)),
label = scales::percent(prop.table(stat(count))))) +  geom_bar() + 
geom_text(stat = "count", position = position_dodge(1),vjust = 1, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent) + coord_flip() +labs(title = "Bookings by Market Segment", x = "MS type",y = "Number of bookings") +
  geom_text(stat = "count", aes(label = ..count..), hjust = 1) 

```
"""

Highest Bookings are with Online TA - 47.3%

"""
# Analysis of Distribution Channel

```{r}

ggplot(data = hData,aes( x = distribution_channel,fill = distribution_channel ,y = prop.table(stat(count)),
label = scales::percent(prop.table(stat(count))))) +  geom_bar() + 
geom_text(stat = "count", position = position_dodge(1),vjust = -0.5, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent)  +labs(title = "Bookings by Distribution Channel", x = "DC type",y = "Number of bookings")

```
"""

Highest Bookings are with TA/TO - 81.98%
followed by Direct  - 12.27%

"""
# Analysis if repeated_guest

```{r}

ggplot(data = hData,aes( x = is_repeated_guest ,y = prop.table(stat(count)),
label = scales::percent(prop.table(stat(count))))) +  geom_bar(fill = 'darkblue') + 
geom_text(stat = "count", position = position_dodge(1),vjust = -0.5, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent)  +labs(title = "Is Repeated Guest", x = "is_repeated_guest",y = "Number of bookings")

```
"""

Repeated guests are negligible with only 3%
Can improve marketing strategy here

"""
# Analysis if previous_cancellations

```{r}

ggplot(data = hData,aes( x = previous_cancellations ,y = prop.table(stat(count)),
label = scales::percent(prop.table(stat(count))))) +  geom_bar(fill = 'darkblue') + 
geom_text(stat = "count", position = position_dodge(1),vjust = -0.5, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent)  +labs(title = "If Previous Cancellations", x = "previous_cancellations",y = "Number of bookings")

```
"""

Only 5% of the booking had previous cancellations

"""
# Analysis by Reserved Room Type

```{r}

ggplot(data = hData,aes( x = reserved_room_type ,y = prop.table(stat(count)),
label = scales::percent(prop.table(stat(count))))) +  geom_bar(fill = 'darkblue') + 
geom_text(stat = "count", position = position_dodge(1),vjust = -0.5, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent)  +labs(title = "Bookings  by Reserved Room Type", x = "Room Type",y = "Number of bookings")

```
"""

72% of bookings are made for Type A
16% for Type D

"""


# Cancelations by Hotel

```{r}
ggplot(data = hData,aes(is_canceled))+ geom_histogram(binwidth = 0.5, fill='darkblue') + facet_wrap(~hotel)

```

# Cancelled average :

```{r echo=TRUE}
canceledRate = hData %>%
  group_by(Hotel = hotel) %>%
  summarise(CanceledAverage = mean(is_canceled)) 
canceledRate = as.data.frame(canceledRate)
canceledRate 
```
"""

Number of cancellations are more in City Hotel

Hotel         CanceledAverage
City Hotel   	0.4172696			
Resort Hotel	0.2776335	

"""


# Cancellation Vs Booking Time 

```{r}
ggplot(data = hData , aes(lead_time)) + geom_histogram(binwidth = 0.8,col='darkblue'
     )+ facet_wrap(~ is_canceled)
```

"""

The cancellations are less when the lead_time is less

"""
