---
title: "RubinaTS"
output: html_document
---


```{r}
hotelData=read.csv('hotel_bookings.csv')
head(hData)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(pROC))
```


#Filter Hotel Data with reservation status and date
```{r}
hdd =  hotelData %>% 
  filter(reservation_status=='Check-Out') %>%
  group_by(reservation_status_date)%>%
  summarise(n=n())

hdd$reservation_status_date = as.Date(hdd$reservation_status_date)

library(lubridate)

hd = hdd %>% group_by(reservation_status_month=floor_date(reservation_status_date, "month")) %>%
  summarise(n=sum(n))


ggplot(hd, aes(reservation_status_month, n)) + geom_line()

```  

#Create TimeSeries for seasonal data
#hs = hotel data seasonal
```{r}

hs = ts(hd$n, start=c(2015, 7), end=c(2017,7), frequency= 12)

```
#Test for stationary time series
```{r}

library(tseries)
adf.test(hs) # p-value < 0.05 indicates the TS is stationary
kpss.test(hs)

```
#See the components of time series
```{r}

components = stl(hs, 'periodic')
plot(components)

```

#Decompose seasonality
```{r}

library(forecast)
seasonplot(hs)

dcomp = decompose(hs)
dcomp$seasonal


```

#Plot to see seasonality
```{r}

plot(forecast(auto.arima(hs)), sub = "Simple plot to forecast")

```
#Forecasts

```{r}
hsforecasts = HoltWinters(hs)
hsforecasts
hsforecasts$fitted
plot(hsforecasts)
library("forecast")
forecast(hsforecasts, h=8)

```
forecast(auto.arima(hsforecasts))

```{r}

library(forecast)
fit <- ets(hs)
fc <- forecast(fit)
plot(fc)
fc

```
> accuracy(fc)
                   ME    RMSE      MAE      MPE     MAPE      MASE      ACF1
Training set 80.61895 510.038 404.3835 1.557431 15.49543 0.7967545 0.1774337


fit_arima <- Arima(hs, order=c(1,1,0), seasonal=list(order=c(1,0,0), period=12))
> accuracy( fit_arima)
                   ME    RMSE      MAE      MPE    MAPE      MASE        ACF1
Training set 22.61754 122.097 101.6285 1.627262 22.3616 0.3061808 -0.05615262


hd = hdd %>% group_by(reservation_status_month=floor_date(reservation_status_date, "month")) %>%
  summarise(n=sum(n))
  
  
=================

#Daily forecast

```{r}
msts = msts(hdd$n, seasonal.periods = c(7,365.25),start =decimal_date(as.Date("2015-07-01")))
plot(msts, main="Daily Admissions", xlab="Year", ylab="Daily Admissions")

tbats = tbats(msts)
plot(tbats, main="Multiple Season Decomposition")

sp = predict(tbats,h=14)
plot(sp, main = "TBATS Forecast", include=14)

```
